---
title: "Twilight Annotation"
format: html
editor: visual
---

## When was twilight?

The workflow recommended by [Lisovski et al.](https://geolocationmanual.vogelwarte.ch/) in their geolocation user manual relies on `TwGeos::preprocessLight()`, which I can't get to work with the Adélie penguin data. So I want to try a different approach using change point analysis.

## Load packages and data

```{r}
#| label: setup

library(lubridate)
library(TwGeos)
library(tidyverse)

move_1947021 <- read_csv("data/gls/1947021.csv")
lightraw <- move_1947021 %>% 
  select(Date = timestamp,
         Light = `gls:light-level`) %>% 
  slice(seq(1, nrow(.), by = 5))

```

## Diagnostic plots

```{r}
#| label: diagnostic-lightimage

# Time offset and calibration location
offset <- -6
lon.calib <- -64.1
lat.calib <- -64.8

# Create a light image plot
lightImage(tagdata = lightraw,
           offset = offset,     
           zlim = c(0, 10))

# Add twilight times for the calibration location
tm <- seq(min(lightraw$Date), max(lightraw$Date), by = "day")
rise <- rep(c(TRUE, FALSE), length(tm))
zenith = 96
c.dat <- data.frame(
  Twilight = SGAT::twilight(rep(tm, each = 2), 
                            lon = lon.calib, lat = lat.calib, 
                            rise = rise, zenith = zenith), 
  Rise = rise
)
tsimageLines(c.dat$Twilight[c.dat$Rise], offset = offset, 
             lwd = 3, col = adjustcolor("orange", alpha.f = 0.5))
tsimageLines(c.dat$Twilight[!c.dat$Rise], offset = offset,
             lwd = 3, col = adjustcolor("orange", alpha.f = 0.5))

```

The light data look very noisy until 2020-02-11, so we'll begin our analysis then.

```{r}
#| label: diagnostic-threshold

sample_begin <- as.POSIXct("2020-03-20", tz = "UTC")
sample_end <- as.POSIXct("2020-03-30", tz = "UTC")
threshold <- 1
filter(lightraw, between(Date, sample_begin, sample_end)) %>% 
  ggplot(aes(x = Date, y = Light)) +
  geom_line() +
  geom_point(aes(color = Light)) +
  geom_hline(yintercept = threshold, linetype = "dashed", color = "red") +
  scale_y_continuous(trans = "pseudo_log", 
                     breaks = 10^(0:5)) +
  scale_color_gradientn(colors = c('black',"purple",'orange'),
                        trans = "pseudo_log", 
                        breaks = 10^(0:5)) +
  theme_classic()

```

This sample of 10 days in late March suggests a light threshold of 1 (red dotted line) is appropriate. Some nights look very clean (e.g., Mar 26) while others are very noisy (e.g., Mar 24).

## Twilight estimates

I'll try to estimate twilight times using changepoint analysis.

```{r}
#| label: twilight-changepoint

light <- lightraw %>% 
  filter(Date >= as.POSIXct("2020-02-11", tz = "UTC")) %>% 
  mutate(is_light = ifelse(Light > threshold, 1, 0))

find_twilights <- function(light, light_thr) {
  is_light <- light > light_thr
  light_cpt <- changepoint::cpt.mean(light,)
}

rise_set <- SGAT::twilight(
  tm = rep(light$Date[1], 2),
  lon = lon.calib,
  lat = lat.calib,
  rise = c(TRUE, FALSE)
)
if (rise_set[1] > rise_set[2]) 
  rise_set[1] <- rise_set[1] + days(1)
noon <- rise_set[1] + (rise_set[2] - rise_set[1]) / 2
light_1day <- light %>% 
  filter(between(Date, noon - hours(13), noon + hours(13)))

light_cpt <- changepoint::cpt.mean(light_1day$is_light, 
                                   method = "PELT", 
                                   Q = 3)
twilights_1day <- data.frame(
  riseset = c("Sunrise", "Sunset"),
  Date = light_1day$Date[light_cpt@cpts[1:2]]
)

ggplot(light_1day, aes(Date, is_light)) + 
  geom_line() +
  geom_vline(aes(xintercept = Date, color = riseset), 
             twilights_1day,
             linewidth = 2) +
  scale_color_manual(values = c("goldenrod", "blue4")) +
  theme_classic() +
  theme(legend.position = "top",
        legend.title = element_blank())
```
